Human Judgement Forecasting of COVID-19 in the UK

### 2023-02-16 manuscript/manuscript.tex
correct errors in manuscript, add citation

\textbf{Background}
Accurate forecasts of infectious diseases can aid public health decision-making. In the past, two studies found ensembles of human judgement forecasts of COVID-19 to show predictive performance comparable to ensembles of computational models, at least when predicting case incidences. We present a follow-up to a study conducted in Germany and Poland and investigate a novel approach to combine human judgement and epidemiological modelling.

\textbf{Methods}
From May 24th to August 16th 2021, we elicited weekly one to four-week-ahead forecasts of cases and deaths from COVID-19 in the UK from a crowd of human forecasters. Forecasts were elicited as part of a tournament, the "UK Crowd Forecasting Challenge", and a median ensemble of all forecasts was submitted to the European Forecast Hub. Participants could use two distinct interfaces: in one, forecasters submitted a predictive distribution directly, in the other forecasters instead submitted a forecast of the effective reproduction number $R_t$. This forecast was automatically mapped to cases and deaths using the renewal equation and a convolution over a delay distribution. Forecasts were scored using the weighted interval score on the original forecasts, as well as after applying the natural logarithm to both forecasts and observations.

\textbf{Results}
The ensemble of human forecasters overall performed comparably to the offical European Forecast Hub ensemble on both cases and deaths, although results were susceptible to changes in details of the evaluation. $R_t$ foreasts performed comparably to direct forecasts on cases, but worse on deaths.

\textbf{Conclusions}
We find more evidence that human judgement can be useful in predicting the large-scale evolution of an infectious disease such as COVID-19. The results of forecast evaluations can change depending on what metrics are chosen and judgement on what does or doesn't constitute a "good" forecast is dependent on the forecast consumer.

@@

The European COVID-19 Forecast Hub \cite{sherrattPredictivePerformanceMultimodel2022a} elicits predictions for various COVID-19 related forecast targets from different research groups every week. Forecasts had to be submitted every Monday 23.59pm GMT. Forecasts are made for incident weekly reported numbers of cases of and deaths from COVID-19 on a national level for various European countries over a one to four week forecast horizon. While forecasts were submitted on Mondays, weeks were defined as epiweeks. ending on a Saturday and starting on Sunday. Forecast horizons were therefore in fact 5, 12, 19 and 26 days. Submissions to the European Forecast Hub follow a quantile-based format with 23 quantiles of each output measure at levels 0.01, 0.025, 0.05, 0.10, 0.15,…, 0.95, 0.975, 0.99.
Every week, forecast submitted to the hub were automatically checked for conformity with the required forecasts and all eligible forecasts combined into different ensembles. Until the 12th of July 2021 the default Hub ensemble shown on all official Forecast Hub visualisiations (https://covid19forecasthub.eu/) was a mean-ensemble (i.e., the $\alpha$-quantile of the ensemble is given by the mean of all submitted $\alpha$-quantiles). From the 29th of July on, the default Forecast Hub ensemble became a median-ensemble. The number of models included in the Forecast Hub ensemble is shown in Figure \ref{fig:num-forecasters} in the SI.

Ground-truth data on daily reported test positive cases and deaths linked to COVID-19 were provided by European Forecast Hub and sourced from the Johns Hopkins University (JHU). Data were subject to reporting artifacts and revisions. All data points are marked as anomalous by the European Forecast Hub if in subsequent updates data is changed by more than 5 percent. In August 2022 JHU switched the data source for their death numbers from "deaths within 28 days of a positive COVID test" to "Deaths with COVID-19 on the death certificate" and revised all their past data to guarantee consistency. The 2021 UK ground truth death data as it was made available through the European Forecast Hub in 2021 is therefore substantially different and on average lower than the data available as of early 2023. The data and revisions are displayed in Figure \ref{fig:plot-data}. If not otherwise stated, all results are derived based on the original data available in 2021.

@@

Forecasts of incident cases and deaths linked to COVID-19 in the UK were elicited from individual participants every week through the web application (https://cmmid-lshtm.shinyapps.io/crowd-forecast/) described in \cite{bosseComparingHumanModelbased2022}. The application is based on \textsf{R} \cite{R} \texttt{shiny} \cite{shiny} and is available as an \textsf{R} package called \texttt{crowdforecastr} \citep{crowdforecastr}.

The web application offered participants two different ways of making a forecast called 'direct' (or 'classical') and 'Rt' forecast. To make a 'direct' forecast (as described in more detail \cite{bosseComparingHumanModelbased2022}), participants selected a predictive distribution (by default a log-normal distribution) and adjusted the median and width of the distribution to change the central estimate and uncertainty. In addition to information about past observations, participants could see various metrics and data sourced from Our World in Data \citep{owidcoronavirus}. Figure \ref{fig:screenshot-classical} shows a screenshot of the forecast interface for direct forecasts.

@@

SAM TO THE RESCUE

In addition, we implemented a second forecasting method ('Rt forecasts'), where we asked participants to make a forecast of the effective reproduction number $R_t$. This forecast was then translated into forecasts of cases using the renewal equation \citep{fraserEstimatingIndividualHousehold2007} as implemented in the \textsf{R} package \texttt{EpiNow2} \citep{epinow2}, APPENDIX INFO?. In order to obtain forecasts for deaths, we convolved predicted infections as implied by the $R_t$ forecast over data-based delay distributions \cite{epinow2, sherrattExploringSurveillanceData2021, abbottEstimatingTimevaryingReproduction2020a} to model the time between infection and report date. As $R_t$-estimates up to two weeks prior to the forecast data were deemed uncertain, we also asked participants to submit an estimate of $R_t$ for the two weeks prior to the current forecast date. Participants were therefore asked to estimate/predict six $R_t$ values, four of them in the future. Upon pressing a button, participants could see a preview of the evolution of cases implied by their current $R_t$ forecast. However, due to the high computational complexity and necessary waiting times, participants could not preview the death forecast implied by their current input for $R_t$. Figure \ref{fig:screenshot-rt} shows a screenshot of the forecast interface for $R_t$ forecasts.

@@

We scored forecasts using the weighted interval score (WIS, formulas given in the SI) %in Section \ref{sec:wis}) 
\cite{bracherEvaluatingEpidemicForecasts2021} and used the empirical coverage of all central 50\% and 90\% prediction intervals to measure probabilistic calibration. The WIS is a proper scoring rule (with smaller values implying better performance). A forecaster, in expectation, optimises her score by providing a predictive distribution $F$ that is equal to the data-generating distribution $G$, and is therefore incentivised to report her true belief. The WIS can be understood as an approximation of the continuous ranked probability score (CRPS) for forecasts in a quantile-based format, which itself represents a generalisation of the absolute error to predictive distributions. The WIS can be decomposed into three separate penalty components: forecast dispersion (i.e. uncertainty of forecasts), over-prediction and under-prediction. Empirical coverage refers to the percentage of observations falling inside any given prediction interval (e.g. the cumulative percentage of observed values that fall inside all 50\% prediction intervals). It represents a measure of probabilistic calibration \citep{gneitingProbabilisticForecastsCalibration2007}. Forecasts were evaluated using the \texttt{scoringutils} \citep{bosseEvaluatingForecastsScoringutils2022} package in \textsf{R}. All code and data, including individual-level forecasting data is available at \url{https://github.com/epiforecasts/uk-crowd-forecasting-challenge}.

\citet{bosseTransformationForecastsEvaluating2023} suggested to transform forecasts and observations using the natural logarithm prior prior to applying the WIS in order to better reflect the exponential nature of the underlying disease process. We therefore also compute WIS values after transforming all forecasts and observations using the function $f\colon x \to \log (x + 1)$. In the following, we refer to WIS scores obtained without a transformation as "scores on the natural scale", and WIS values obtained after log-transforming forecasts and observations as "scores on the log scale".

@@

For two-week-ahead-forecasts of incident cases, the Forecast Hub ensemble outperformed all crowd forecasting ensembles as measured by the WIS on the natural scale (see Figure \ref{fig:performance} and Table \ref{tab:scores}). On the natural scale, the WIS as a measure of the absolute distance between forecast and observation increases or decreases with the magnitude of the forecast target \cite{bosseTransformationForecastsEvaluating2023, bracherEvaluatingEpidemicForecasts2021}. Average scores are therefore dominated by performance around the peak when cases were highest, in particular by forecasts made on the 19th of July for the 31st of July (see Figure \ref{fig:forecasts-scores}). $R_t$ forecasts tended to be higher than both the direct forecasts and the Forecast Hub ensemble, especially around the peak, and therefore showed the poorest performance in terms of WIS on the natural scale (see Figure \ref{fig:performance} and Table \ref{tab:scores}). When looking at WIS values on the log scale,
%(that is, when log-transforming forecasts and observations prior to applying the WIS), 
scores are more equally distributed across the study period and more weight is given to forecasts in June and July which underpredicted the extent to which case number would rise. On the log scale, all human forecast ensembles performed slightly better than the Forecast Hub ensemble. Similarly to what was observed previously for case forecasts \citep{bosseComparingHumanModelbased2022, sherrattPredictivePerformanceMultimodel2022a}, all forecasting approaches exhibited underdispersion, meaning that forecasts on average were too narrow and not uncertain enough, with empirical coverage below nominal coverage. Underdispersion was most severe for the $R_t$ forecasts.

@@

In terms of the WIS on the natural scale, direct human forecasts performed best, while $R_t$ forecasts performed significantly worse than all other approaches (see Figure \ref{fig:performance} and Table \ref{tab:scores}). From June to the end of July, $R_t$ forecasts overpredicted deaths and were noticeable higher than other forecasts, whereas in August $R_t$ forecasts underpredicted deaths and were substantially lower than other forecasts (see Figure \ref{fig:forecasts-scores}). On the natural scale, the highest scores were concentrated towards the end of the study period, where death incidences were highest, while on the log scale, scores were more evenly distributed across the study period. On the log scale, direct human forecasts, the combined crowd ensemble and the Forecast Hub ensemble showed comparable performance, while $R_t$ forecasts performed again significantly worse. 
%Note that the picture changes when looking at the revised data which was corrected upwards compared to the original data(DO WE WANT TO DO THAT? IS THAT MEANINGFUL?) (see Figure \ref{fig:data-revisions}). For the revised data, direct forecasts performed significantly worse than both the Forecast Hub ensemble and the $R_t$ forecasts on the log scale, while still performing better than the $R_t$ forecasts on the natural scale (see Table \ref{tab:scores-revised}). 
%For revised data: On the natural scale, the largest scores occurred at the end of the study period when deaths were highest. On the log scale, however, underprediction in June and July when death incidences were lower, received more weight and direct crowd forecast outperformed $R_t$ forecasts. 
$R_t$ forecasts show strong underdispersion with empirical coverage of the 50\% prediction intervals being close to zero and smaller than 50\% for the 90\% prediction intervals (see Figure \ref{fig:performance} and Table \ref{tab:scores}). The Forecast Hub ensemble was overdispersed with empirical coverage exceeding nominal coverage, and for the direct human forecasts, empirical coverage was close to nominal coverage.

@@

In this paper, we presented a follow-up study to \cite{bosseComparingHumanModelbased2022}, analysing human judgement forecasts of cases of and deaths from COVID-19 in the United Kingdom submitted to the European COVID-19 Forecast Hub between the 24th of May and the 16th of August 2021. Human judgement forecasts were generated using two different forecasting approaches, a) direct forecasts of cases and deahts and b) forecasts of the effective reproduction number $R_t$ which were mapped to cases using the renewal equation and convolved over a delay distribution to obtain deaths. We found that human judgement forecasts of case numbers, regardless of how they were obtained, showed performance broadly comparable to the ensemble of all forecasts submitted to the European Forecast Hub. For deaths, performance of direct human forecasts was comparable to that of the Forecast Hub ensemble, while $R_t$ forecasts performed significantly worse. Model rankings were not always robust, but sometimes changed depending on how forecasts are evaluated. For example, model rankings could differ when focusing on a different forecast horizon, or analysing WIS values on the natural scale vs. on the log scale (i.e. after log-transforming both forecasts and observations).

@@

In their original study conducted in Germany and Poland, \citet{bosseComparingHumanModelbased2022} found that humans outperformed an ensemble of computational models when predicting cases, but not when predicting deaths. They hypothesised that computational models might have an advantage over human forecsters when predicting deaths, benefitting from the ability to model the delays and exact epidemiological relationships between different leading and lagged indicators. \citet{mcandrewChimericForecastingCombining2022} similarly found that humans performed comparably to an ensemble of computational models for cases, but not for predictions of deaths of COVID-19. We could not replicate these findings, and found that in terms of the WIS on the natural scale, direct human forecasts were slightly worse than the Forecast Hub ensemble for cases, and slightly better for deaths (although results for greater forecast horizons differ). Looking at scores on the log scale suggests overall comparable performance of the Hub ensemble and the our direct human forecasts on both cases and deaths. Relatively good performance of direct human judgement forecasts for death forecasts compared to previous studies may be related to the rise of the Delta variant in the UK starting in the beginning of May. The Delta variant had a higher case fatality ratio (CFR) than previous variants CITATION. Humans may have found it easier than computational models to adapt to the fact that the relationship between cases and deaths continuously changed during the study period.

>>>>

\paragraph{Background}

Accurate forecasts of infectious diseases can aid public health decision-making. In the past, two studies found ensembles of human judgement forecasts of COVID-19 to show predictive performance comparable to ensembles of computational models, at least when predicting case incidences. We present a follow-up to a study conducted in Germany and Poland and investigate a novel approach to combine human judgement and epidemiological modelling.

\paragraph{Methods}

From May 24th to August 16th 2021, we elicited weekly one to four-week-ahead forecasts of cases and deaths from COVID-19 in the UK from a crowd of human forecasters. Forecasts were elicited as part of a tournament, the "UK Crowd Forecasting Challenge", and a median ensemble of all forecasts was submitted to the European Forecast Hub. Participants could use two distinct interfaces: in one, forecasters submitted a predictive distribution directly, in the other forecasters instead submitted a forecast of the effective reproduction number $R_t$. This forecast was automatically mapped to cases and deaths using the renewal equation and a convolution over a delay distribution. Forecasts were scored using the weighted interval score on the original forecasts, as well as after applying the natural logarithm to both forecasts and observations.

\paragraph{Results}

The ensemble of human forecasters overall performed comparably to the offical European Forecast Hub ensemble on both cases and deaths, although results were susceptible to changes in details of the evaluation. $R_t$ foreasts performed comparably to direct forecasts ona cases, but worse on deaths. Self-identified experts did not outperform non-experts.

\paragraph{Conclusions}

Human judgement can be useful in predicting the large-scale evolution of an infectious disease such as COVID-19. The results of forecast evaluations can change depending on what metrics are chosen and judgement on what does or doesn't constitute a "good" forecast is dependent on the forecast consumer.

@@

The European COVID-19 Forecast Hub \cite{sherrattPredictivePerformanceMultimodel2022a} elicits predictions for various COVID-19 related forecast targets from different research groups every week. Forecasts had to be submitted every Monday 23.59pm GMT. Forecasts are made for incident weekly reported numbers of cases of and deaths from COVID-19 on a national level for various European countries over a one to four week forecast horizon. While forecasts were submitted on Mondays, weeks were defined as epiweeks, ending on a Saturday and starting on Sunday. Forecast horizons were therefore in fact 5, 12, 19 and 26 days. Submissions to the European Forecast Hub follow a quantile-based format with 23 quantiles of each output measure at levels 0.01, 0.025, 0.05, 0.10, 0.15,…, 0.95, 0.975, 0.99.
Every week, forecasts submitted to the hub were automatically checked for conformity with the required format and all eligible forecasts combined into different ensembles. Until the 12th of July 2021 the default Hub ensemble shown on all official Forecast Hub visualisiations (https://covid19forecasthub.eu/) was a mean-ensemble (i.e., the $\alpha$-quantile of the ensemble is given by the mean of all submitted $\alpha$-quantiles). From the 29th of July on, the default Forecast Hub ensemble became a median-ensemble. The number of models included in the Forecast Hub ensemble is shown in Figure \ref{fig:num-forecasters} in the SI.

Ground-truth data on daily reported test positive cases and deaths linked to COVID-19 were provided by the European Forecast Hub and sourced from the Johns Hopkins University (JHU). Data were subject to reporting artifacts and revisions. All data points are marked as anomalous by the European Forecast Hub if in subsequent updates data is changed by more than 5 percent. In August 2022 JHU switched the data source for their death numbers from "deaths within 28 days of a positive COVID test" to "Deaths with COVID-19 on the death certificate" and revised all their past data to guarantee consistency. The 2021 UK ground truth death data as it was made available through the European Forecast Hub in 2021 is therefore substantially different and on average lower than the data available as of early 2023. The data and revisions are displayed in Figure \ref{fig:plot-data}. All results presented here were derived based on the original data available in 2021.

@@

Forecasts of incident cases and deaths linked to COVID-19 in the UK were elicited from individual participants every week through a web application (https://cmmid-lshtm.shinyapps.io/crowd-forecast/) described in \cite{bosseComparingHumanModelbased2022}. The application is based on \textsf{R} \cite{R} \texttt{shiny} \cite{shiny} and is available as an \textsf{R} package called \texttt{crowdforecastr} \citep{crowdforecastr}.

@@

The web application offered participants two different ways of making a forecast, called 'direct' (or 'classical') and 'Rt' forecast. To make a 'direct' forecast (as described in more detail in \cite{bosseComparingHumanModelbased2022}), participants selected a predictive distribution (by default a log-normal distribution) and adjusted the median and width of the distribution to change the central estimate and uncertainty. In addition to information about past observations, participants could see various metrics and data sourced from Our World in Data \citep{owidcoronavirus}. Figure \ref{fig:screenshot-classical} shows a screenshot of the forecast interface for direct forecasts.

\textbf{SAM TO THE RESCUE}

In addition to the 'direct' forecasts, we implemented a second forecasting method ('Rt forecasts'), where we asked participants to make a forecast of the effective reproduction number $R_t$. This forecast was then translated into forecasts of cases using the renewal equation \citep{fraserEstimatingIndividualHousehold2007} as implemented in the \textsf{R} package \texttt{EpiNow2} \citep{epinow2} (see details in the SI). In order to obtain forecasts for deaths, we convolved predicted infections as implied by the $R_t$ forecast over data-based delay distributions \cite{epinow2, sherrattExploringSurveillanceData2021, abbottEstimatingTimevaryingReproduction2020a} to model the time between infection and report date. As $R_t$-estimates up to two weeks prior to the forecast data were deemed uncertain, we also asked participants to submit an estimate of $R_t$ for the two weeks prior to the current forecast date. Participants were therefore asked to estimate/predict six $R_t$ values, four of them in the future. Upon pressing a button, participants could see a preview of the evolution of cases implied by their current $R_t$ forecast. However, due to the high computational complexity and necessary waiting times, participants could not preview the death forecast implied by their current input for $R_t$. Figure \ref{fig:screenshot-rt} shows a screenshot of the forecast interface for $R_t$ forecasts.

@@

We scored forecasts using the weighted interval score (WIS, formulas given in the SI) %in Section \ref{sec:wis}) 
\cite{bracherEvaluatingEpidemicForecasts2021} and used the empirical coverage of all central 50\% and 90\% prediction intervals to measure probabilistic calibration \citep{gneitingProbabilisticForecastsCalibration2007}. The WIS is a proper scoring rule (with smaller values implying better performance). A forecaster, in expectation, optimises her score by providing a predictive distribution $F$ that is equal to the data-generating distribution $G$, and is therefore incentivised to report her true belief. The WIS can be understood as an approximation of the continuous ranked probability score (CRPS) for forecasts in a quantile-based format, which itself represents a generalisation of the absolute error to predictive distributions. The WIS can be decomposed into three separate penalty components: forecast dispersion (i.e. uncertainty of forecasts), over-prediction and under-prediction. Empirical coverage refers to the percentage of observations falling inside any given central prediction interval (e.g. the cumulative percentage of observed values that fall inside all central 50\% prediction intervals). Forecasts were evaluated using the \texttt{scoringutils} \citep{bosseEvaluatingForecastsScoringutils2022} package in \textsf{R}. All code and data, including individual-level forecasting data is available at \url{https://github.com/epiforecasts/uk-crowd-forecasting-challenge}.

\citet{bosseTransformationForecastsEvaluating2023} suggested to transform forecasts and observations using the natural logarithm prior to applying the WIS in order to better reflect the exponential nature of the underlying disease process. We therefore also compute WIS values after transforming all forecasts and observations using the function $f\colon x \to \log (x + 1)$. In the following, we refer to WIS scores obtained without a transformation as "scores on the natural scale", and WIS values obtained after log-transforming forecasts and observations as "scores on the log scale".

@@

For two-week-ahead-forecasts of incident cases, the Forecast Hub ensemble outperformed all crowd forecasting ensembles as measured by the WIS on the natural scale (see Figure \ref{fig:performance} and Table \ref{tab:scores}). The difference between the Hub ensemble and the human judgement forecasts steadily increased with increasing forecast horizon. On the natural scale, the WIS as a measure of the absolute distance between forecast and observation increases or decreases with the magnitude of the forecast target \cite{bosseTransformationForecastsEvaluating2023, bracherEvaluatingEpidemicForecasts2021}. Average scores are therefore dominated by performance around the peak when cases were highest, in particular by forecasts made on the 19th of July for the 31st of July (see Figure \ref{fig:forecasts-scores}). $R_t$ forecasts tended to be higher than both the direct forecasts and the Forecast Hub ensemble, especially around the peak, and therefore showed the poorest performance in terms of WIS on the natural scale (see Figure \ref{fig:performance} and Table \ref{tab:scores}). When looking at WIS values on the log scale,
%(that is, when log-transforming forecasts and observations prior to applying the WIS), 
scores are more equally distributed across the study period and more weight is given to forecasts in June and July which underpredicted the extent to which case number would rise. On the log scale, all human forecast ensembles performed slightly better than the Forecast Hub ensemble. For a four week forecast horizon, however, the Forecast Hub ensemble performed slightly better than human forecasts. Similarly to what was observed previously for case forecasts \citep{bosseComparingHumanModelbased2022, sherrattPredictivePerformanceMultimodel2022a}, all forecasting approaches exhibited underdispersion, meaning that forecasts on average were too narrow and not uncertain enough, with empirical coverage below nominal coverage. Underdispersion was most severe for the $R_t$ forecasts.

@@

In terms of the WIS on the natural scale, direct human forecasts performed best at a two-week forecast horizon, while $R_t$ forecasts performed significantly worse than all other approaches (see Figure \ref{fig:performance} and Table \ref{tab:scores}). From June to the end of July, $R_t$ forecasts overpredicted deaths and were noticeable higher than other forecasts, whereas in August, $R_t$ forecasts underpredicted deaths and were substantially lower than other forecasts (see Figure \ref{fig:forecasts-scores}). For four-week-ahead death forecasts, the Hub ensemble performed slightly better than other approaches. 
On the natural scale, the highest scores were concentrated towards the end of the study period, where death incidences were highest, while on the log scale, scores were more evenly distributed across the study period. On the log scale, direct human forecasts, the combined crowd ensemble and the Forecast Hub ensemble showed comparable performance, while $R_t$ forecasts performed again significantly worse. 
%Note that the picture changes when looking at the revised data which was corrected upwards compared to the original data(DO WE WANT TO DO THAT? IS THAT MEANINGFUL?) (see Figure \ref{fig:data-revisions}). For the revised data, direct forecasts performed significantly worse than both the Forecast Hub ensemble and the $R_t$ forecasts on the log scale, while still performing better than the $R_t$ forecasts on the natural scale (see Table \ref{tab:scores-revised}). 
%For revised data: On the natural scale, the largest scores occurred at the end of the study period when deaths were highest. On the log scale, however, underprediction in June and July when death incidences were lower, received more weight and direct crowd forecast outperformed $R_t$ forecasts. 
$R_t$ forecasts show strong underdispersion with empirical coverage of the 50\% prediction intervals being close to zero and smaller than 50\% for the 90\% prediction intervals (see Figure \ref{fig:performance} and Table \ref{tab:scores}). The Forecast Hub ensemble was overdispersed with empirical coverage exceeding nominal coverage, and for the direct human forecasts, empirical coverage was close to nominal coverage.

@@

In this paper, we presented a follow-up study to \cite{bosseComparingHumanModelbased2022}, analysing human judgement forecasts of cases of and deaths from COVID-19 in the United Kingdom submitted to the European COVID-19 Forecast Hub between the 24th of May and the 16th of August 2021. Human judgement forecasts were generated using two different forecasting approaches, a) direct forecasts of cases and deaths and b) forecasts of the effective reproduction number $R_t$, which were mapped to cases using the renewal equation and convolved over a delay distribution to obtain deaths. We found that human judgement forecasts of case numbers, regardless of how they were obtained, showed performance broadly comparable to the ensemble of all forecasts submitted to the European Forecast Hub. For deaths, performance of direct human forecasts was comparable to that of the Forecast Hub ensemble, while $R_t$ forecasts performed significantly worse. Model rankings were not always robust, but sometimes changed depending on how forecasts are evaluated. For example, model rankings could differ when focusing on a different forecast horizon, or analysing WIS values on the natural scale vs. on the log scale (i.e. after log-transforming both forecasts and observations).

@@

In their original study conducted in Germany and Poland, \citet{bosseComparingHumanModelbased2022} found that humans outperformed an ensemble of computational models when predicting cases, but not when predicting deaths. They hypothesised that computational models might have an advantage over human forecsters when predicting deaths, benefitting from the ability to model the delays and exact epidemiological relationships between different leading and lagged indicators. \citet{mcandrewChimericForecastingCombining2022} similarly found that humans performed comparably to an ensemble of computational models for cases, but not for predictions of deaths of COVID-19. We could not replicate these findings, and found that in terms of the WIS on the natural scale, direct human forecasts were slightly worse than the Forecast Hub ensemble for cases, and slightly better for deaths (although results for greater forecast horizons differed). Looking at scores on the log scale suggests overall comparable performance of the Hub ensemble and the our direct human forecasts on both cases and deaths. Relatively good performance of direct human judgement forecasts for death forecasts compared to previous studies may be related to the rise of the Delta variant in the UK starting in the beginning of May. The Delta variant had a higher case fatality ratio (CFR) than previous variants \cite{linDiseaseSeverityClinical2021}. Humans may have found it easier than computational models to adapt to the fact that the relationship between cases and deaths continuously changed during the study period.
